Language and vision in conceptual processing: Multilevel analysis and statistical power

### 2022-10-23 manuscript/manuscript.tex
23102022

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

@@

This manuscript is a draft. Correspondence can be addressed to Pablo Bernabeu on \href{mailto:pcbernabeu@gmail.com}{\nolinkurl{pcbernabeu@gmail.com}}. All materials needed to inspect and reproduce the present research---comprising documentation, data, analysis and results---are available at \url{http://doi.org/10.17605/OSF.IO/UERYQ}.

@@

\abstract{%
Research has suggested that conceptual processing depends on both language-based and vision-based information. We tested this interplay at three levels of the experimental structure: individuals, words and tasks. To this end, we drew on three existing data sets that were composed of large samples, and implemented the paradigms of semantic priming, semantic decision and lexical decision. We extended these data sets with language-based and vision-based measures, and analysed their interactions with participants' vocabulary size and gender, and with presentation speed in the semantic priming study. We performed the analysis using mixed-effects models that included a comprehensive array of fixed effects---including covariates---and random effects. First, language-based information was found to be more important than vision-based information. Second, in the semantic priming study---whose task required distinguishing between words and nonwords---, both language-based and vision-based information were more influential when words were presented faster. Third, a `task-relevance advantage' was identified in higher-vocabulary participants. Specifically, in lexical decision, higher-vocabulary participants were more sensitive than lower-vocabulary participants to language-based information. In contrast, in semantic decision, higher-vocabulary participants were more sensitive to word concreteness. Fourth, a human-based measure of visual information was found to be superior to a computational measure. These findings support the interplay between language and vision in conceptual processing, and demonstrate the influence of measurement instruments on the results. Last, we estimated the sample size required to reliably investigate various effects. We found that 300 participants were sufficient to examine the effect of language-based information contained in words, whereas more than 1,000 participants were necessary for the effect of vision-based information and for the interactions of both former variables with vocabulary size, gender and presentation speed. In conclusion, this power analysis reveals the need to increase sample sizes in research on perceptual simulation and individual differences.
}

@@

Over the last two decades, research in cognitive science has suggested that conceptual processing depends on both language and embodiment systems. That is, understanding words involves---on the one hand---lexical and semantic associations of an amodal kind, and---on the other hand---modality-specific associations within perceptual, motor, affective and social domains (Barsalou et al., 2008; Connell, 2019; Davis \& Yee, 2021; Khatin-Zadeh et al., 2021; Roads \& Love, 2020; Vigliocco et al., 2009). Studies addressing these systems have found that the language system is overall more prevalent in word processing, producing larger effect sizes (Banks et al., 2021; Kiela \& Bottou, 2014; Lam et al., 2015; Louwerse et al., 2015; Pecher et al., 1998; Petilli et al., 2021). More intricately, the roles of language and embodiment are modulated by the characteristics of individuals, words and tasks. For instance, people's individual experience with language is associated with differential effects relating to phonological, lexical and semantic features of words (Jared \& O'Donnell, 2017; Pexman \& Yap, 2018; Yap et al., 2009, 2012, 2017). Similarly, physical expertise and perceptual biases are associated with differences in the mental simulation of meaning (Beilock et al., 2008; Calvo-Merino et al., 2005; Vukovic \& Williams, 2015). Furthermore, the embodiment system is especially suited for the processing of concrete concepts---e.g., \emph{red}, \emph{building} (C. R. Jones et al., 2022; Kousta et al., 2011; Ponari, Norbury, Rotaru, et al., 2018; cf. Borghi et al., 2022). Embodied information also becomes more important in the following conditions: (I) later in the time courses of word recognition (Bernabeu et al., 2017; Louwerse \& Hutchinson, 2012; cf. Petilli et al., 2021) and property generation (Santos et al., 2011; Simmons et al., 2008), (II) when participants produce slower responses (Louwerse \& Connell, 2011), and (III) in tasks that elicit deeper semantic processing---e.g., semantic decision, as opposed to lexical decision (Ostarek \& Huettig, 2017; Petilli et al., 2021).

@@

We revisit three larger-than-average studies to investigate the interplay between language and embodiment in conceptual processing. We devote a study to each of the three original studies. Thus, Study 1 is centred on Hutchison et al. (2013) and uses the semantic priming paradigm. Study 2 is centred on Pexman et al. (2017) and uses the semantic decision paradigm. Study 3 is centred on Balota et al. (2007) and uses the lexical decision paradigm. Each of these central studies contained measures of participants' vocabulary size and gender. Furthermore, the core data sets were expanded by adding variables that captured the language-based information in words (Mandera et al., 2017; Wingfield \& Connell, 2022) and the vision-based information in words (Lynott et al., 2020; Petilli et al., 2021)---the latter being used to represent the embodiment system. One of the key questions we investigated using this array of variables was whether individual differences in vocabulary and gender modulated participants' sensivity to the language-based and vision-based information in words. Alongside the effects of interest, several covariates were included in the models to allow a rigorous analysis (Sassenhagen \& Alday, 2016). These covariates comprised measures of general cognition and lexical characteristics of the stimulus words. Last, in each study, we performed a statistical power analysis to help estimate the sample size needed to investigate a variety of effects in future studies.

@@

Studies have operationalised the language system at the word level using measures that capture the relationships among words without explicitly drawing on any sensory or affective modalities. Two main types of linguistic measures exist: those based on text corpora---dubbed \emph{word co-occurrence} measures (Bullinaria \& Levy, 2007; Petilli et al., 2021; Wingfield \& Connell, 2022)---and those based on associations collected from human participants---dubbed \emph{word association} measures (De Deyne et al., 2016, 2019). Notwithstanding the interrelation between word co-occurrence and word association (Planchuelo et al., 2022), co-occurrence is more purely linguistic, whereas association indirectly captures more of the sensory and affective meaning of words (De Deyne et al., 2021).

@@

In Study 1 (semantic priming) and Study 2 (semantic decision), co-occurrence measures were used to represent the language system at the word level. Specifically, in Study 1, this measure was called \texttt{language-based\ similarity}, and it was based on the degree of text-based co-occurrence between the prime word and the target word in each trial (Mandera et al., 2017). In Study 2, the measure was called \texttt{word\ co-occurrence}, and it was based on the degree of text-based co-occurrence between each stimulus word and the words `abstract' and `concrete' (Wingfield \& Connell, 2022). In Study 3 (lexical decision), a co-occurrence measure could not be used, as the co-occurrence of words in consecutive trials could not be calculated due to the high frequency of nonword trials throughout the lexical decision task. Therefore, a single-word measure had to be used instead. Word frequency was used as it was the lexical variable, among five, that had the largest effect (see \protect\hyperlink{appendix-A-lexical-covariates}{\underline{Appendix A}}).

@@

We hypothesised that word co-occurrence, word frequency and vocabulary size would all have facilitatory effects on participants' performance, with higher values leading to shorter RTs (Pexman \& Yap, 2018; Wingfield \& Connell, 2022; Yap et al., 2009).

@@

In previous studies, the embodiment system has been represented at the word level by perceptual, motor, affective or social variables (Fernandino et al., 2022; Vigliocco et al., 2009; Wang et al., 2021). For instance, the perceptual modalities have often corresponded to the five Aristotelian senses---vision, hearing, touch, taste and smell (Bernabeu et al., 2017, 2021; Louwerse \& Connell, 2011)---and, less often, to interoception (Connell et al., 2018). Yet, out of all these senses, vision has been most frequently used in research (e.g., Bottini et al., 2021; De Deyne et al., 2021; Pearson \& Kosslyn, 2015; Petilli et al., 2021; Yee et al., 2012). The hegemony of vision is likely due to the central position of this sense in the human brain (Reilly et al., 2020) as well as in several languages (Bernabeu, 2018; Chen et al., 2019; Lynott et al., 2020; Miceli et al., 2021; Morucci et al., 2019; Roque et al., 2015; Speed \& Brybaert, 2021; Speed \& Majid, 2020; Vergallito et al., 2020; Winter et al., 2018; Zhong et al., 2022). In the present study, we focussed on vision alone both because of its prominence in cognition and because we had planned to use the present research to determine the sample size of a subsequent study that focusses on vision (indeed, the present study grew out of a statistical power analysis).

@@

Experimental data in psycholinguistics can be divided into various levels, such as individuals, words and tasks. The simultaneous examination of a theory across several levels is expected to enhance our understanding of the theory (Ostarek \& Bottini, 2021)---for instance, by revealing the distribution of explanatory power within and across these levels. Several studies have probed more than one level---for instance, word level and individual level (Aujla, 2021; Lim et al., 2020; Pexman \& Yap, 2018; Yap et al., 2009), or word level and task level (Al-Azary et al., 2022; Connell \& Lynott, 2013, 2014; Ostarek \& Huettig, 2019; Petilli et al., 2021). This multilevel approach is complementary to a different line of research that aims to test the causality of various sources of information in conceptual processing, such as language (Ponari, Norbury, Rotaru, et al., 2018), perception (Stasenko et al., 2014) and action (Speed et al., 2017).

@@

The word level is concerned with the lexical and semantic information in words (e.g., De Deyne et al., 2021; Lam et al., 2015; Lund et al., 1995; Lund \& Burgess, 1996; Lynott et al., 2020; Mandera et al., 2017; Petilli et al., 2021; Pexman et al., 2017; Santos et al., 2011; Wingfield \& Connell, 2022). The word-level variables of interest in this study are language-based and vision-based information (both described above). The covariates are lexical variables and word concreteness. The lexical covariates were selected in each study out of the same five variables (see \protect\hyperlink{covariates}{\underline{Covariates}} section below).

@@

Beyond task-level variables, there is an additional source of task-related information across the three studies---namely, the experimental paradigm used in each study (i.e., semantic priming, semantic decision and lexical decision). Indeed, it is possible to examine how the effects vary across these paradigms (see Wingfield \& Connell, 2022). This comparison, however, must be considered cautiously due to the existence of other non-trivial differences across these studies, such as the numbers of observations. With this caveat noted, the tasks used across these studies likely elicit varying degrees of semantic depth, as ordered below (see Balota \& Lorch, 1986; Barsalou et al., 2008; Becker et al., 1997; de Wit \& Kinoshita, 2015; Joordens \& Becker, 1997; Lam et al., 2015; Muraki \& Pexman, 2021; Ostarek \& Huettig, 2017; Versace et al., 2021; Wingfield \& Connell, 2022).

@@

The central objective of the present studies is the simultaneous investigation of language-based and vision-based information, along with the interactions between each of those and vocabulary size, gender and presentation speed (i.e., SOA). Previous studies have examined subsets of these effects using the same data sets we are using (Balota et al., 2007; Petilli et al., 2021; Pexman et al., 2017; Pexman \& Yap, 2018; Wingfield \& Connell, 2022; Yap et al., 2012, 2017). Out of these studies, only Petilli et al. (2021) investigated \emph{both} language and vision. However, in contrast to our present study, Petilli et al.~did not examine the role of vocabulary size or any other individual differences, instead collapsing the data across participants.

@@

The analytical method was broadly similar across the three studies. Below, we present the commonalities in the statistical analysis and in the power analysis.

@@

Lexical covariates were selected in every study out of the same five variables, which had been used as covariates in Wingfield and Connell (2022; also see Petilli et al., 2021). They comprised: number of letters (i.e., orthographic length), word frequency, number of syllables (both the latter from Balota et al., 2007), orthographic Levenshtein distance (Yarkoni et al., 2008) and phonological Levenshtein distance (Suárez et al., 2011; Yap \& Balota, 2009). The selection among these candidates was performed because some of them were highly intercorrelated---i.e., \(r\) \textgreater{} .70 (Dormann et al., 2013; Harrison et al., 2018). The correlations and the selection models are available in \protect\hyperlink{appendix-A-lexical-covariates}{\underline{Appendix A}}.

@@

With regard to random effects, participants and stimuli were \emph{crossed} in the three studies. That is, each participant was presented with a subset of the stimuli. Conversely, each word was presented to a subset of participants. Therefore, linear mixed-effects models were implemented. These models included a maximal random-effects structure, with by-participant and by-item random intercepts, and the appropriate random slopes for all effects of interest (Barr et al., 2013). In the semantic priming study, the items were prime--target pairs, whereas in the semantic decision and lexical decision studies, the items were individual words. In the case of interactions, random slopes were included only when the interacting variables varied within the same unit (Brauer \& Curtin, 2018)---e.g., an interaction of two variables varying within participants (only present in Study 1). Where required due to convergence warnings, random slopes for covariates were removed, as inspired by Remedy 11 from Brauer and Curtin (2018). In this regard, whereas Brauer and Curtin (2018) contemplate the removal of random slopes for covariates only when the covariates are not interacting with any effects of interest, we removed random slopes for covariates even if they interacted with effects of interest because these interactions were covariates themselves.

@@

\(P\) values were calculated using the Kenward-Roger approximation for degrees of freedom (Luke, 2017) in the R package `lmerTest', Version 3.1-3 (Kuznetsova et al., 2017). The latter package in turn used `lme4', Version 1.1-26 (Bates et al., 2015; Bates et al., 2021). To facilitate the convergence of the models, the maximum number of iterations was set to 1 million. Diagnostics regarding convergence and normality are provided in \protect\hyperlink{appendix-B-frequentist-analysis-diagnostics}{\underline{Appendix B}}. Effects that were non-significant or very small are best interpreted by considering their 95\% confidence intervals, and the 95\% credible intervals (i.e., the Bayesian counterpart) where available (Cumming, 2014).

@@

{\centering \includegraphics[width=1\linewidth]{/Users/pablodejuanbernabeu/Library/CloudStorage/OneDrive-LancasterUniversity/semanticpriming-semanticdecision-lexicaldecision/bayesian_priors/plots/bayesian_priors}

@@

Power analyses require setting an effect size for each effect. Often, it is difficult to determine the effect size, as the amount and the scope of relevant research are usually finite and biased (Albers \& Lakens, 2018; Gelman \& Carlin, 2014; Kumle et al., 2021). In some power analyses, the original effect sizes from previous studies have been adopted without any modification (e.g., Pacini \& Barnard, 2021; Villalonga et al., 2021). In contrast, some authors have opted to reduce the previous effect sizes to account for two intervening factors. First, publication bias and insufficient statistical power cause published effect sizes to be inflated (Brysbaert, 2019; Loken \& Gelman, 2017; Open Science Collaboration, 2015; Vasishth, Mertzen, et al., 2018; Vasishth \& Gelman, 2021). Second, over the course of the research, a variety of circumstances could create differences between the planned study and the studies that were used in the power analysis. Some of these differences could be foreseeable---for instance, if they are due to a limitation in the literature available for the power analysis---, whereas other differences might be unforeseeable and could go unnoticed (Barsalou, 2019). Reducing the effect size in the power analysis leads to an increase of the sample size of the planned study (Brysbaert \& Stevens, 2018; Green \& MacLeod, 2016; Hoenig \& Heisey, 2001). The reduced effect size---sometimes dubbed the smallest effect size of interest---is often set with a degree of arbitrariness. In previous studies, Fleur et al. (2020) applied a reduction of 1/8 (i.e., 12.5\%), whereas Kumle et al. (2021) applied a 15\% reduction. In the present study, a reduction of 20\% was applied to every effect in the power analysis. By comparison with the power analyses reviewed in this paragraph, the present reduction will lead to a more conservative estimate of required sample sizes. However, after considering the precedents of small samples and publication bias reviewed above, a 20\% reduction is arguably a reasonable safeguard. Indeed, a posteriori, the results of our power analyses suggested that the 20\% reduction had not been excessive, as some of the effects examined were detectable with small sample sizes.

@@

In some studies, the association between prime and target words has been investigated in terms of related versus unrelated pairs (Lam et al., 2015; Pecher et al., 1998; Trumpp et al., 2013) and---in other studies---in terms of first- and second-order relationships (Hutchison et al., 2013). In contrast to these categorical associations, a third set of studies have measured the association between the prime and the target words using continuous estimates of text-based similarity (Günther et al., 2016a, 2016b; Hutchison et al., 2008; M. N. Jones et al., 2006; Lund et al., 1995; Lund \& Burgess, 1996; Mandera et al., 2017; McDonald \& Brew, 2002; Padó \& Lapata, 2007; Petilli et al., 2021; Wingfield \& Connell, 2022). In one of these studies, Mandera et al. (2017) found that computational measures of similarity outperformed human-based associations at explaining language-based priming.

@@

\begin{itemize}
\item
  \texttt{Vocabulary\ size}. The test used by Hutchison et al. (2013) comprised a synonym test, an antonym test, and an analogy test, all three extracted from the Woodcock--Johnson III diagnostic reading battery (Woodcock et al., 2001). We operationalised the vocabulary measure as the mean score across the three tasks per participant.
\item
  \texttt{Language-based\ similarity}. This measure was calculated using a semantic space from Mandera et al. (2017), which the authors found to be the second-best predictor (\(R\)\textsuperscript{2} = .465) of the semantic priming effect in the lexical decision task of Hutchison et al. (2013) (we could not use the best semantic space, \(R\)\textsuperscript{2} = .471, owing to computational limitations). The second-best semantic space (see first row in Table 5 in Mandera et al., 2017) was based on lemmas from a subtitle corpus, and was processed using a Continuous Bag Of Words model. It had 300 dimensions and a window size of six words.
\item
  \texttt{Stimulus\ onset\ asynchrony\ (SOA)}. Following Brauer and Curtin (2018), the categories of this factor were recoded as follows: 200 ms = -0.5, 1,200 ms = 0.5.
\end{itemize}

@@

>>>>

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

@@

This manuscript is a draft. Correspondence can be addressed to Pablo Bernabeu on \href{mailto:pcbernabeu@gmail.com}{\nolinkurl{pcbernabeu@gmail.com}}. All the materials needed to inspect and reproduce the present research---comprising documentation, data, analysis and results---are available at \url{http://doi.org/10.17605/OSF.IO/UERYQ}.

@@

\abstract{%
Research has suggested that conceptual processing depends on both language-based and vision-based information. We tested this interplay at three levels of the experimental structure: individuals, words and tasks. To this end, we drew on three existing, large data sets that implemented the paradigms of semantic priming, semantic decision and lexical decision. We extended these data sets with measures of language-based and vision-based information, and analysed how the latter variables interacted with participants' vocabulary size and gender, and also with presentation speed in the semantic priming study. We performed the analysis using mixed-effects models that included a comprehensive array of fixed effects---including covariates---and random effects. First, we found that language-based information was found to be more important than vision-based information. Second, in the semantic priming study---whose task required distinguishing between words and nonwords---, both language-based and vision-based information were more influential when words were presented faster. Third, a `task-relevance advantage' was identified in higher-vocabulary participants. Specifically, in lexical decision, higher-vocabulary participants were more sensitive to language-based information than lower-vocabulary participants. In contrast, in semantic decision, higher-vocabulary participants were more sensitive to word concreteness. Fourth, we demonstrated the influence of the analytical method on the results. These findings support the interplay between language and vision in conceptual processing, and demonstrate the influence of measurement instruments on the results. Last, we estimated the sample size required to reliably investigate various effects. We found that 300 participants were sufficient to examine the effect of language-based information contained in words, whereas more than 1,000 participants were necessary for the effect of vision-based information and for the interactions of both former variables with vocabulary size, gender and presentation speed. In conclusion, this power analysis reveals the need to increase sample sizes for research on perceptual simulation and individual differences.
}

@@

Over the last two decades, research in the cognitive sciences has suggested that conceptual processing depends on both language and embodiment systems. That is, understanding words involves---on the one hand---lexical and semantic associations of an amodal kind, and---on the other hand---modality-specific associations within perceptual, motor, affective and social domains (Barsalou et al., 2008; Connell, 2019; Davis \& Yee, 2021; Khatin-Zadeh et al., 2021; Vigliocco et al., 2009). Studies addressing these systems have found that the language system is overall more prevalent in word processing, producing larger effect sizes (Banks et al., 2021; Kiela \& Bottou, 2014; Lam et al., 2015; Louwerse et al., 2015; Pecher et al., 1998; Petilli et al., 2021). More intricately, the roles of language and embodiment are modulated by the characteristics of individuals, words and tasks. For instance, people's individual experience with language is associated with differential effects relating to phonological, lexical and semantic features of words (Jared \& O'Donnell, 2017; Pexman \& Yap, 2018; Yap et al., 2009, 2012, 2017). Similarly, physical expertise and perceptual biases are associated with differences in the mental simulation of meaning (Beilock et al., 2008; Calvo-Merino et al., 2005; Vukovic \& Williams, 2015). Furthermore, the embodiment system is especially suited for the processing of concrete concepts---e.g., \emph{red}, \emph{building} (C. R. Jones et al., 2022; Kousta et al., 2011; Ponari, Norbury, Rotaru, et al., 2018; cf. Borghi et al., 2022). Embodied information also becomes more important in the following conditions: (I) later in the time courses of word recognition (Bernabeu et al., 2017; Louwerse \& Hutchinson, 2012; cf. Petilli et al., 2021) and property generation (Santos et al., 2011; Simmons et al., 2008), (II) when participants produce slower responses (Louwerse \& Connell, 2011), and (III) in tasks that elicit deeper semantic processing---e.g., semantic decision, as opposed to lexical decision (Ostarek \& Huettig, 2017; Petilli et al., 2021). Last, research in computational linguistics has provided further support for the complementarity of language and embodied information, by revealing increased predictive performance when models are provided with perceptual information on top of text-based information (Frome et al., 2013; Roads \& Love, 2020).

@@

We revisit three larger-than-average studies to investigate the interplay between language and embodiment in conceptual processing. We devote a study to each of the three original studies. Thus, Study 1 is centred on Hutchison et al. (2013) and uses the semantic priming paradigm. Study 2 is centred on Pexman et al. (2017) and uses the semantic decision paradigm. Study 3 is centred on Balota et al. (2007) and uses the lexical decision paradigm. Each of these central studies contained measures of participants' vocabulary size and gender. Furthermore, the core data sets were expanded by adding variables that captured the language-based information in words (Mandera et al., 2017; Wingfield \& Connell, 2022a) and the vision-based information in words (Lynott et al., 2020; Petilli et al., 2021)---the latter being used to represent the embodiment system. One of the key questions we investigated using this array of variables was whether individual differences in vocabulary and gender modulated participants' sensivity to the language-based and vision-based information in words. Alongside the effects of interest, several covariates were included in the models to allow a rigorous analysis (Sassenhagen \& Alday, 2016). These covariates comprised measures of general cognition and lexical characteristics of the stimulus words. Last, in each study, we performed a statistical power analysis to help estimate the sample size needed to investigate a variety of effects in future studies.

@@

Studies have operationalised the language system at the word level using measures that capture the relationships among words without explicitly drawing on any sensory or affective modalities. Two main types of linguistic measures exist: those based on text corpora---dubbed \emph{word co-occurrence} measures (Bullinaria \& Levy, 2007; Petilli et al., 2021; Wingfield \& Connell, 2022a)---and those based on associations collected from human participants---dubbed \emph{word association} measures (De Deyne et al., 2016, 2019). Notwithstanding the interrelation between word co-occurrence and word association (Planchuelo et al., 2022), co-occurrence is more purely linguistic, whereas association indirectly captures more of the sensory and affective meaning of words (De Deyne et al., 2021).

@@

In Study 1 (semantic priming) and Study 2 (semantic decision), co-occurrence measures were used to represent the language system at the word level. Specifically, in Study 1, this measure was called \texttt{language-based\ similarity}, and it was based on the degree of text-based co-occurrence between the prime word and the target word in each trial (Mandera et al., 2017). In Study 2, the measure was called \texttt{word\ co-occurrence}, and it was based on the degree of text-based co-occurrence between each stimulus word and the words `abstract' and `concrete' (Wingfield \& Connell, 2022a). In Study 3 (lexical decision), a co-occurrence measure could not be used, as the co-occurrence of words in consecutive trials could not be calculated due to the high frequency of nonword trials throughout the lexical decision task. Therefore, a single-word measure had to be used instead. Word frequency was used as it was the lexical variable, among five, that had the largest effect (see \protect\hyperlink{appendix-A-lexical-covariates}{\underline{Appendix A}}).

@@

We hypothesised that word co-occurrence, word frequency and vocabulary size would all have facilitatory effects on participants' performance, with higher values leading to shorter RTs (Pexman \& Yap, 2018; Wingfield \& Connell, 2022a; Yap et al., 2009).

@@

In previous studies, the embodiment system has been represented at the word level by perceptual, motor, affective or social variables (Fernandino et al., 2022; Vigliocco et al., 2009; Wang et al., 2021). For instance, the perceptual modalities have often corresponded to the five Aristotelian senses---vision, hearing, touch, taste and smell (Bernabeu et al., 2017, 2021; Louwerse \& Connell, 2011)---and, less often, to interoception (Connell et al., 2018). Yet, out of all these domains, vision has been most frequently used in research (e.g., Bottini et al., 2021; De Deyne et al., 2021; Pearson \& Kosslyn, 2015; Petilli et al., 2021; Yee et al., 2012). The hegemony of vision is likely due to the central position of vision in the human brain (Reilly et al., 2020) as well as in several languages (Bernabeu, 2018; Chen et al., 2019; Lynott et al., 2020; Miceli et al., 2021; Morucci et al., 2019; Roque et al., 2015; Speed \& Brybaert, 2021; Speed \& Majid, 2020; Vergallito et al., 2020; Winter et al., 2018; Zhong et al., 2022). In the present study, we focussed on vision alone due to three reasons. First, we wanted to use a single variable to represent sensorimotor information, just as a single variable would be used to represent linguistic information. Using a single variable for each system facilitates the analysis of interactions with other variables. Second, vision is very prominent in cognition, as we just reviewed. Third, we had planned to use the present research to determine the sample size of a subsequent study that focusses on vision (indeed, the present study grew out of a statistical power analysis).

@@

Experimental data in psycholinguistics can be divided into various levels, such as individuals, words and tasks. The simultaneous examination of a theory across several levels is expected to enhance our understanding of the theory (Ostarek \& Bottini, 2021)---for instance, by revealing the distribution of explanatory power (that is, effect size) within and across these levels. Several studies have probed more than one level---for instance, word level and individual level (Aujla, 2021; Lim et al., 2020; Pexman \& Yap, 2018; Yap et al., 2009), or word level and task level (Al-Azary et al., 2022; Connell \& Lynott, 2013, 2014; Ostarek \& Huettig, 2019; Petilli et al., 2021). This multilevel approach is complementary to a different line of research that aims to test the causality of various sources of information in conceptual processing, such as language (Ponari, Norbury, Rotaru, et al., 2018), perception (Stasenko et al., 2014) and action (Speed et al., 2017).

@@

The word level is concerned with the lexical and semantic information in words (e.g., De Deyne et al., 2021; Lam et al., 2015; Lund et al., 1995; Lund \& Burgess, 1996; Lynott et al., 2020; Mandera et al., 2017; Petilli et al., 2021; Pexman et al., 2017; Santos et al., 2011; Wingfield \& Connell, 2022a). The word-level variables of interest in this study are language-based and vision-based information (both described above). The covariates are lexical variables and word concreteness. The lexical covariates were selected in each study out of the same five variables (see \protect\hyperlink{covariates}{\underline{Covariates}} section below).

@@

Beyond task-level variables, there is an additional source of task-related information across the three studies---namely, the experimental paradigm used in each study (i.e., semantic priming, semantic decision and lexical decision). Indeed, it is possible to examine how the effects vary across these paradigms (see Wingfield \& Connell, 2022a). This comparison, however, must be considered cautiously due to the existence of other non-trivial differences across these studies, such as the numbers of observations. With this caveat noted, the tasks used across these studies likely elicit varying degrees of semantic depth, as ordered below (see Balota \& Lorch, 1986; Barsalou et al., 2008; Becker et al., 1997; de Wit \& Kinoshita, 2015; Joordens \& Becker, 1997; Lam et al., 2015; Muraki \& Pexman, 2021; Ostarek \& Huettig, 2017; Versace et al., 2021; Wingfield \& Connell, 2022a).

@@

The central objective of the present studies is the simultaneous investigation of language-based and vision-based information, along with the interactions between each of those and vocabulary size, gender and presentation speed (i.e., SOA). Previous studies have examined subsets of these effects using the same data sets we are using (Balota et al., 2007; Petilli et al., 2021; Pexman et al., 2017; Pexman \& Yap, 2018; Wingfield \& Connell, 2022a; Yap et al., 2012, 2017). Out of these studies, only Petilli et al. (2021) investigated \emph{both} language and vision. However, in contrast to our present study, Petilli et al.~did not examine the role of vocabulary size or any other individual differences, instead collapsing the data across participants.

@@

The analytical method was broadly similar across the three studies. Below, we present the commonalities in the statistical analysis and in the power analysis. Several R packages from the `tidyverse' (Wickham et al., 2019) were used.

@@

Lexical covariates were selected in every study out of the same five variables, which had been used as covariates in Wingfield and Connell (2022a; also see Petilli et al., 2021). They comprised: number of letters (i.e., orthographic length), word frequency, number of syllables (both the latter from Balota et al., 2007), orthographic Levenshtein distance (Yarkoni et al., 2008) and phonological Levenshtein distance (Suárez et al., 2011; Yap \& Balota, 2009). The selection among these candidates was performed because some of them were highly intercorrelated---i.e., \(r\) \textgreater{} .70 (Dormann et al., 2013; Harrison et al., 2018). The correlations and the selection models are available in \protect\hyperlink{appendix-A-lexical-covariates}{\underline{Appendix A}}.

@@

With regard to random effects, participants and stimuli were \emph{crossed} in the three studies. That is, each participant was presented with a subset of the stimuli. Conversely, each word was presented to a subset of participants. Therefore, linear mixed-effects models were implemented. These models included a maximal random-effects structure, with by-participant and by-item random intercepts, and the appropriate random slopes for all effects of interest (Barr et al., 2013; Brauer \& Curtin, 2018; Singmann \& Kellen, 2019). Random effects---especially random slopes---constrain the analytical space by \emph{claiming} their share of variance. As a result, that variance cannot be taken by the fixed effects. In the semantic priming study, the items were prime--target pairs, whereas in the semantic decision and lexical decision studies, the items were individual words. In the case of interactions, random slopes were included only when the interacting variables varied within the same unit (Brauer \& Curtin, 2018)---e.g., an interaction of two variables varying within participants (only present in Study 1). Where required due to convergence warnings, random slopes for covariates were removed, as inspired by Remedy 11 from Brauer and Curtin (2018). In this regard, whereas Brauer and Curtin (2018) contemplate the removal of random slopes for covariates only when the covariates are not interacting with any effects of interest, we removed random slopes for covariates even if they interacted with effects of interest because these interactions were covariates themselves.

@@

\(P\) values were calculated using the Kenward-Roger approximation for degrees of freedom (Luke, 2017) in the R package `lmerTest', Version 3.1-3 (Kuznetsova et al., 2017). The latter package in turn used `lme4', Version 1.1-26 (Bates et al., 2015; Bates et al., 2021). To facilitate the convergence of the models, the maximum number of iterations was set to 1 million. Diagnostics regarding convergence and normality are provided in \protect\hyperlink{appendix-B-frequentist-analysis-diagnostics}{\underline{Appendix B}}. Those effects that are non-significant or very small are best interpreted by considering the confidence intervals and the credible intervals (Cumming, 2014).

The R package `GGally' (Schloerke et al., 2021) was used to create correlation plots, whereas the package `sjPlot' (Lüdecke, 2021) was used for interaction plots.

@@

{\centering \includegraphics[width=1\linewidth]{C:/Users/Pablo/Documents/semanticpriming-semanticdecision-lexicaldecision/bayesian_priors/plots/bayesian_priors}

@@

Power analyses require setting an effect size for each effect. Often, it is difficult to determine the effect size, as the amount and the scope of relevant research are usually finite and biased (Albers \& Lakens, 2018; Gelman \& Carlin, 2014; Kumle et al., 2021). In some power analyses, the original effect sizes from previous studies have been adopted without any modification (e.g., Pacini \& Barnard, 2021; Villalonga et al., 2021). In contrast, some authors have opted to reduce the previous effect sizes to account for two intervening factors. First, publication bias and insufficient statistical power cause published effect sizes to be inflated (Brysbaert, 2019; Loken \& Gelman, 2017; Open Science Collaboration, 2015; Vasishth, Mertzen, et al., 2018; Vasishth \& Gelman, 2021). Second, over the course of the research, a variety of circumstances could create differences between the planned study and the studies that were used in the power analysis. Some of these differences could be foreseeable---for instance, if they are due to a limitation in the literature available for the power analysis---, whereas other differences might be unforeseeable and could go unnoticed (Barsalou, 2019; Noah et al., 2018). Reducing the effect size in the power analysis leads to an increase of the sample size of the planned study (Brysbaert \& Stevens, 2018; Green \& MacLeod, 2016; Hoenig \& Heisey, 2001). The reduced effect size---sometimes dubbed the smallest effect size of interest---is often set with a degree of arbitrariness. In previous studies, Fleur et al. (2020) applied a reduction of 1/8 (i.e., 12.5\%), whereas Kumle et al. (2021) applied a 15\% reduction. In the present study, a reduction of 20\% was applied to every effect in the power analysis. By comparison with the power analyses reviewed in this paragraph, the present reduction will lead to a more conservative estimate of required sample sizes. However, after considering the precedents of small samples and publication bias reviewed above, a 20\% reduction is arguably a reasonable safeguard. Indeed, a posteriori, the results of our power analyses suggested that the 20\% reduction had not been excessive, as some of the effects examined were detectable with small sample sizes.

@@

In some studies, the association between prime and target words has been investigated in terms of related versus unrelated pairs (Lam et al., 2015; Pecher et al., 1998; Trumpp et al., 2013) and---in other studies---in terms of first- and second-order relationships (Hutchison et al., 2013). In contrast to these categorical associations, a third set of studies have measured the association between the prime and the target words using continuous estimates of text-based similarity (Günther et al., 2016a, 2016b; Hutchison et al., 2008; M. N. Jones et al., 2006; Lund et al., 1995; Lund \& Burgess, 1996; Mandera et al., 2017; McDonald \& Brew, 2002; Padó \& Lapata, 2007; Petilli et al., 2021; Wingfield \& Connell, 2022a). In one of these studies, Mandera et al. (2017) found that computational measures of similarity outperformed human-based associations at explaining language-based priming.

@@

\begin{itemize}
\item
  \texttt{Vocabulary\ size}. The test used by Hutchison et al. (2013) comprised a synonym test, an antonym test, and an analogy test, all three extracted from the Woodcock--Johnson III diagnostic reading battery (Woodcock et al., 2001). We operationalised the vocabulary measure as the mean score across the three tasks per participant.
\item
  \texttt{Language-based\ similarity}. This measure was calculated using a semantic space from Mandera et al. (2017), which the authors found to be the second-best predictor (\(R\)\textsuperscript{2} = .465) of the semantic priming effect in the lexical decision task of Hutchison et al. (2013) (we could not use the best semantic space, \(R\)\textsuperscript{2} = .471, owing to computational limitations). The second-best semantic space (see first row in Table 5 in Mandera et al., 2017) was based on lemmas from a subtitle corpus, and was processed using a Continuous Bag Of Words model. It had 300 dimensions and a window size of six words. The R package `LSAfun' (Günther et al., 2015) was used to import this variable.\footnote{Despite the name of the package, the measure we used was not based on Latent Semantic Analysis.}
\item
  \texttt{Stimulus\ onset\ asynchrony\ (SOA)}. Following Brauer and Curtin (2018), the categories of this factor were recoded as follows: 200 ms = -0.5, 1,200 ms = 0.5.
\end{itemize}

@@

\leavevmode\vadjust pre{\hypertarget{ref-R-LSAfun}{}}%
Günther, F., Dudschig, C., \& Kaup, B. (2015). LSAfun: An r package for computations based on latent semantic analysis. \emph{Behavior Research Methods}, \emph{47}(4), 930--944. \url{https://doi.org/10.3758/s13428-014-0529-0}
